comgen-outline

### 2017-10-31 paper/comgen-paper.Rmd
move bda details to appendix

### Bayesian data analysis

In order to integrate the results of our prior measurement into richer models of endorsement, we first try to accurately model the prior data by itself.
We do this using a Bayesian mixture model, which assumes that the data generated by participants for each property (e.g., \textsc{lay eggs}) comes from one of two underlying distributions. 
One of these distributions is fixed to represent kinds of animals who *do not have* an internal mechanism that could give rise to the property (e.g., \textsc{lions} and \textsc{lay eggs}).
The second distribution is flexible and represents kinds of animals who *do have* such an internal mechanism; this distribution is flexible and can represent different shapes for different properties (e.g., the differences in feature-probability judgments for \textsc{lay eggs} vs. \text{has wings}).

Formally, we assume the prior data (analzyed independently for each property) was generated from one of two distributions: a distribution corresponding to those kinds with a stable causal mechanism that *could* give rise to the property ($\mathcal{D}_{stable}$) and a "transient cause" distribution corresponding to those kinds without a stable mechanism ($\mathcal{D}_{transient}$).
The "transient" distribution intuitively corresponds to accidental causes of the feature (e.g., a lion, who through some genetic mutation, reproduces by laying eggs).
We model this distribution as a Beta distribuition that heavily favors probabilities near 0: $\text{Beta}(\gamma = 0.01, \delta = 100)$.^[
Note that we use the noncanonical mean $\gamma$ and concentration $\xi$ (or, inverse-variance) parameterization of the Beta distribution rather than the canonical shape (or pseudocount) parameterization for ease of posterior inference. The shape parameterization can be recovered using: $\alpha = \gamma \cdot \xi; \beta = (1 - \gamma) \cdot \xi$.
]
The "stable" distribution is modeled as a Beta distribution with unknown parameters $\text{Beta}(\gamma, \xi)$.^[
Because the Beta distribution is not defined at the points 0 and 1, we add $\epsilon$ to the 0 responses and round 1 to 0.99.
Adjusting 1 to $1- \epsilon$ leads to improper inferences for this model, as $1 - \epsilon$ is only likely under a highly right-skewed distribution; treating 1 as $1- \epsilon$ disproportionately influences the shape of $\mathcal{D}_{stable}$ relative to other responses.
This problem does not appear for 0 being adjusted to $\epsilon$ because the "transient" distribution already expects such low values.
Similar results can be obtained by rounding 0 to 0.01. 
Alternatively, the "transient" distribution could be defined as a Delta distribution at 0, and 0 responses could remain in their raw form.
]
Finally, we assume that these two components combine with mixture weighting $\phi$ such that the data we observe is $$P(d) = \phi\cdot \text{Beta} (d \mid \gamma, \xi) + (1 -  \phi) \cdot \text{Beta}(d \mid \gamma = 0.01, \xi = 100) $$.
<!-- The unstructured alternative hypothesis assumes the data was generated by a single component in the prior distribution with unknown parameteres $P(d) = \text{Beta} (d \mid \gamma, \xi)$. -->
We put the following priors over the latent parameters of the model:
\begin{eqnarray}
\phi_i & \sim & \text{Uniform}(0, 1) \\
\gamma_i & \sim & \text{Uniform}(0, 1) \\
\xi_i & \sim & \text{Uniform}(0, 100)
\end{eqnarray}
where $i$ ranges over the different properties (e.g., *lays eggs*, *carries malaria*).

@@

@@

To learn about the credible values of the parameters, we ran separate MCMC chains for each item, collecting 75,000 samples, removing the first 25,000 for burn-in.
To see how well the mixture model fits the feature-probability prior data, we use the inferred parameters to generate new data.
The data generated from the model's posterior is called the *posterior predictive distribution* and is an important step in model criticism. 
If the model is a good representation of the data, the posterior predictive data will align with the observed experimental data. 
We construct a posterior predictive distribution by "forward sampling" the model.^[
This can be described by the following algorithm: First, flip a coin weighted by $\phi$.
If it comes up heads, we then sample from the "stable" component: $\text{Beta}(\gamma, \xi)$.
If it comes up tails, we sample from the "transient" component: $\text{Beta}(0.01, 100)$.
We do this many times using the posterior distibution to generate a distribution over predicted feature-probability ratings.
]

@@

For both fixed-threshold and uncertain-threshold models, the endorsement model $E$ (Eq. \ref{eq:S1}) decides whether or not to assert the generalization conditioned on a particular referent-category feature-probability $h'$^[
Rather than trying to communicate a particular $h$, the speaker could plausibly try to communicate her full belief distribution over $h$ to the listener. 
This corresponds to the more general form of the model described in a footnote in the model section. 
For these data, the predictions of these two models are nearly identical.
]
(e.g., given the percentage of robins that lay eggs, does it want to say "Robins lay eggs"?).
We use the empirically estimated referent-category feature-probability (the same as used in the baseline regression model) as the $h'$ in the endorser $E$ model.
The empirically measured priors from Expt. 1b are used as the literal interpreter's feature-probability prior $P(h)$.
The $E$ endorsement model is then fully specified (i.e., can generate predictions) after setting the optimality parameter $\lambda_1$ (in Eq. \ref{eq:S1}).

```{r genericsModelDiagram, fig.cap="Graphical model corresponding to the fully Bayesian data analysis of the RSA model. The feature-probability prior data $d^{pr}_{f}$ is assumed to be generated from the mixture model validated in Expt. 1a. The referent-category feature-probability $d^{pr}_{k,f}$ is generated from a latent feature-probability $h_{k,f}$. The fixed-threshold (control model) or uncertain-threshold RSA model takes the feature-probability prior $P_f(h)$, the referent-category feature-probability $h$, and a single speaker optimality parameter $\\lambda$ as input. This overall structured is repeated (except $\\lambda$) for each of the unique properties $f$ and categories $k$ that correspond to the generic sentences in our stimulus set. Note that RSA corresponds to a probabilistic *function* and not a *random variable* that is standard in graphical model notation; RSA cannot be represented by a graphical model because it has recursion. "}
knitr::include_graphics("figs/generics_bayesnetmodel.pdf", dpi = 108)
```

To fit this pair of endorsement models, we incorporate them into the Bayesian data analytic model of the feature-probability prior data (Expt. 1b) to create a single, joint-inference model where the optimality parameter $\lambda_1$ is inferred jointly with all the other latent parameters of the full model (the referent feature-probability $h'_{k, f}$ for each category $k$ and property $f$ and the parameters of the feature-probability priors $P_f(h)$ for each property $f$) using data from Expt. 1a \& b (see Figure \ref{fig:genericsModelDiagram}).
For the parameters of the feature-probability prior distributions, we use the same priors described in Expt. 1b; for the speaker optimality parameter, we use a prior with a range consistent with previous literature using the same model class: $\lambda_1 \sim \text{Uniform}(0,5)$.
We learn about the \emph{a posteriori} credible values of the joint inference models by collecting samples from 3 MCMC chains of 100,000 iterations removing the first 50,000 iterations for burn-in, using an incrementalized version of the Metropolis-Hastings algorithm [@Ritchie2016].

@@

>>>>

@@

In order to integrate the results of our prior measurement into richer models of endorsement, we first try to accurately model the prior data by itself.
We do this using a Bayesian mixture model, which assumes that the data generated by participants for each property (e.g., \textsc{lay eggs}) comes from one of two underlying distributions. 
One of these distributions is fixed to represent kinds of animals who *do not have* an internal mechanism that could give rise to the property (e.g., \textsc{lions} and \textsc{lay eggs}).
The second distribution is flexible and represents kinds of animals who *do have* such an internal mechanism; this distribution is flexible and can represent different shapes for different properties (e.g., the differences in feature-probability judgments for \textsc{lay eggs} vs. \text{has wings}).
For model implementation and inference details, see Appendix.

@@

@@

For both fixed-threshold and uncertain-threshold models, the endorsement model $E$ (Eq. \ref{eq:S1}) decides whether or not to assert the generalization conditioned on a particular referent-category feature-probability $h'$^[
  Rather than trying to communicate a particular $h$, the speaker could plausibly try to communicate her full belief distribution over $h$ to the listener. 
  This corresponds to the more general form of the model described in a footnote in the model section. 
  For these data, the predictions of these two models are nearly identical.
]
(e.g., given the percentage of robins that lay eggs, does it want to say "Robins lay eggs"?).
We use the empirically estimated referent-category feature-probability (the same as used in the baseline regression model) as the $h'$ in the endorser $E$ model.
The empirically measured priors from Expt. 1b are used as the literal interpreter's feature-probability prior $P(h)$.
The $E$ endorsement model is then fully specified (i.e., can generate predictions) after setting the optimality parameter $\lambda_1$ (in Eq. \ref{eq:S1}).
To fit this pair of endorsement models, we incorporate them into the Bayesian data analytic model of the feature-probability prior data (Expt. 1b) to create a single, joint-inference model where the optimality parameter $\lambda_1$ is inferred jointly with all the other latent parameters of the full model (the referent feature-probability $h'_{k, f}$ for each category $k$ and property $f$ and the parameters of the feature-probability priors $P_f(h)$ for each property $f$) using data from Expt. 1a \& b (see Appendix and Figure \ref{fig:genericsModelDiagram}).

@@

```{r child = 'appendix-bda.Rmd'}
```
